---
title: "STAT/MATH 495: Problem Set 09"
author: "WRITE YOUR NAME(S) HERE"
date: "2017-11-07"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE, warning = FALSE
  )
set.seed(76)

library(tidyverse)
```



# Collaboration

Please indicate who you collaborated with on this assignment: 



# Question 1: Run k-means

```{r}
observations_1 <- read_csv("data/observations_1.csv")
observations_2 <- read_csv("data/observations_2.csv")

# Set observations to be one of two datasets
observations <- observations_2

# Fit model for k=2
k <- 2
k_means_results <- kmeans(observations, centers=k)
clusters <- k_means_results$cluster
cluster_centers <- k_means_results$centers

# Add cluster results to observations. Note we convert to factor since cluster
# ID's should be treated as categorical
observations$cluster <- as.factor(clusters)

# Add cluster ID's to cluster_centers
cluster_centers <- cluster_centers %>% 
  as_tibble() %>% 
  mutate(cluster=as.factor(1:k))

ggplot(NULL, aes(x=x1, y=x2, col=cluster)) +
  geom_point(data=observations) +
  geom_point(data=cluster_centers, size=5)
```

**Questions**:

1. Run KMC 10 times on `observations_1` and comment on the consistency of the
results.
1. Speculate on the root cause of any consistency or inconsistency in the
results.
1. Run KMC 10 times on `observations_2` and comment on the consistentcy of the
results.
1. Speculate on the root cause of any consistency or inconsistency in the
results.

**Answers**:

1.
1.
1.
1.



# Bonus question: Code your own

Read ISLR page 388 Algorithm 10.1 and implement k-means clustering from scratch.
Don't worry about doing it for general $k$; keep it simple and do it for $k=2$
specifically. Apply it to `observations_2` from above.

```{r}
# Hint:
library(proxy)
A <- data_frame(
  x1 = c(0, 0.5, 0.75, 1),
  x2 = c(0, 0.5, 0.75, 1)
)
B <- data_frame(
  x1 = c(1, 0),
  x2 = c(1, 0)
)
distance_matrix <- proxy::dist(x=A, y=B)
distance_matrix
apply(distance_matrix, 1, which.min)
```
